Paper

### 2020-07-14 Paper.Rmd
Updates to the paper, and generating a standardized patient/sample database for metadata

```{r demo}

cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% 
  dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (dates)'=paste(min(start.year), max(end.year), sep='-'),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  #'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  #'Total Range endoscopies'=paste(range(total.endos), collapse='-'),
  #'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Median total endoscopies (IQR)'=paste0(median(total.endos), ' (', paste(quantile(total.endos, prob=c(0.25,0.75)), collapse='-'), ')'),
  'Median age at diagnosis (IQR)'=paste0(median(age.diagnosis), ' (', paste(quantile(age.diagnosis, prob=c(0.25,0.75)), collapse='-'), ')'),
  'Mean Length (not recorded)'=paste0(paste(round(mean(M,na.rm=T), 1),'+/-',round(sd(M,na.rm=T),1)), ' (', length( which(is.na(M) )),')'),
  'Gender F:M'=paste(table(gender),collapse=':'),
  'Smoking N:Y (not recorded)'=paste0(paste(table(smoking), collapse=':'), ' (', length(which(is.na(smoking) )), ')')
) %>% left_join(patient.info %>% group_by(Status, Pathology) %>% tally %>% spread(Pathology,n), by='Status')

@@

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(dir, 'Analysis/models_5e6_all/50kb', sep='/')

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

@@

#melted$brks = as.integer(factor(cut(melted$value, breaks = quantile(melted$value, probs=c(seq(0,1,0.05))), include.lowest = T)))
p = ggplot(melted, aes(Var2, Var1, fill = value)) + geom_tile() +
  scale_fill_gradient2() + 
  theme(axis.text = element_blank()) + labs(x='Feature', y='sWGS Sample')
leg = get_legend(p)
ggsave('plots/samples.png', plot = p + theme(legend.position = 'none'), width=3, height=4, units='in')

@@

riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]

@@

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=6, fig.width=10}
chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

@@

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))

@@

## NDBE
nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))

globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot +  labs(y='mean adj. seg.', title='All samples') + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)

gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)

@@

tt = wilcox.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

mp = model.performance(performance.at.1se, coefs, folds, splits )
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling(full_width = F)
```

@@

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
rm(performance.at.1se, coefs, mp)
```

@@

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
rm(performance.at.1se, coefs, mp)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p
ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, performance.at.1se)

@@

p = ggplot(features, aes(label,hazards, fill=coef)) + geom_col() + 
  scale_fill_gradient2() + coord_flip() + theme_classic() + 
  theme(axis.ticks.y = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), legend.position = 'none') + labs(x='', y='', title='Effects') 
ggsave('plots/effects.png', plot=p, width = 3, height=1.5, units='in')

cvRR = features
qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

p = globalCN.plot + geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
```

@@

file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

@@

p = ggplot(dplyr::select(pg.samp, Status, Samplename, RR), aes(reorder(Samplename, -RR), 1, fill=RR)) + geom_tile() + 
  #scale_fill_gradientn(colors=myPal) +
  scale_fill_gradient2(midpoint = 5, low = scales::muted('blue'), high = scales::muted('red')) + 
  scale_y_continuous(expand=c(0,0)) + 
  coord_flip() + labs(y='', x='') + theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.position = 'none') 
ggsave('plots/per_sample_RR.png', plot = p, width=0.5, height=3, units='in')

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
```

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

@@

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T, 0.3)

@@

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
```

@@

```{r loonolgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

#length(performance.at.1se)
#length(unique(pg.sampNOLGD$Patient))

#perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOLGD$Patient))

@@

ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
```

@@

#pROC::coords(roc, 0.3, input='threshold', transpose=T)

@@

blues = RColorBrewer::brewer.pal(5,'Greys')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
```

@@

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )

hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD only', incS = T), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% 
  dplyr::summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID) )

@@

multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo), threshold = 0.3) + labs(title='Endoscopy prediction ROCs')
```

@@

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtM, 'best', transpose=T)
rocPtM$model = 'Max per Patient'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

>>>>

```{r demo}
cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% 
  dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (dates)'=paste(min(start.year), max(end.year), sep='-'),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  #'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  #'Total Range endoscopies'=paste(range(total.endos), collapse='-'),
  #'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Median total endoscopies (IQR)'=paste0(median(total.endos), ' (', paste(quantile(total.endos, prob=c(0.25,0.75)), collapse='-'), ')'),
  'Median age at diagnosis (IQR)'=paste0(median(age.diagnosis), ' (', paste(quantile(age.diagnosis, prob=c(0.25,0.75)), collapse='-'), ')'),
  'Mean Length (not recorded)'=paste0(paste(round(mean(M,na.rm=T), 1),'+/-',round(sd(M,na.rm=T),1)), ' (', length( which(is.na(M) )),')'),
  'Gender F:M'=paste(table(gender),collapse=':'),
  'Smoking N:Y (not recorded)'=paste0(paste(table(smoking), collapse=':'), ' (', length(which(is.na(smoking) )), ')')
) %>% left_join(patient.info %>% group_by(Status, Pathology) %>% tally %>% spread(Pathology,n), by='Status')

@@

rm(cohortSummary, sample.age, fit)
```

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(dir, 'Analysis/models_5e6_all/50kb', sep='/') 
file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

@@

p = ggplot(melted, aes(Var2, Var1, fill = value)) + geom_tile() +
  scale_fill_gradient2() + 
  theme(axis.text = element_blank()) + labs(x='Feature', y='sWGS Sample')
leg = get_legend(p)
ggsave('plots/samples.png', plot = p + theme(legend.position = 'none'), width=3, height=4, units='in')

@@

@@

rm(p,leg, melted)
```

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=6, fig.width=10}
rm(residuals)

chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

@@

n.status <- c('Progressor'=paste0('P\n(n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('NP\n(n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))

@@

gp = globalCN.plot + labs(x='Chromosome', y='Relative CN', title='') + 
  theme(legend.position='none', axis.text.y=element_text(size=8), strip.text.x=element_text(size=8), strip.text.y=element_text(size=9), title=element_text(face = 'bold'), axis.title=element_text(face="plain")) + 
  scale_y_continuous(breaks = c(-20,0,20) )
ggsave(filename='plots/globalCN.png', plot=gp, width=10, height=2.2, units='in', dpi=300)

## NDBE
nb.status <- c('Progressor'=paste0('P\n(n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
              'Non-Progressor'=paste0('NP\n(n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))

globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples')

gbe = globalBECN.plot + labs(y='Relative CN',x='Chromosome', title="") +
  theme(legend.position='none', axis.text.y=element_text(size=8), strip.text.x=element_text(size=8), strip.text.y=element_text(size=9), title=element_text(face = 'bold'), axis.title=element_text(face="plain")) + 
  scale_y_continuous(breaks = c(-10,0,10)) 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=10, height=2.2, units='in', dpi=300)

rm(arm.loc, cnLegend)
```

@@

tt = wilcox.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)

cx.labels = complexity.score %>% group_by(Status) %>% tally() %>% spread(Status, n)

complexity.score %>% dplyr::rename('Complexity Score' = 'cx') %>% dplyr::select(-Patient, -Pathology) %>% write_tsv('plots/cx_score.tsv')

pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8, width=0.3) +
  scale_x_discrete(breaks = unique(complexity.score$Status), 
                   labels = c('Non-Progressor' = paste0('Non-Progressor\n(n=',cx.labels$`Non-Progressor`,')'),'Progressor' = paste0('Progressor\n(n=',cx.labels$`Progressor`,')'))) + 
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rm(cx,complexity.score, tt, cx.labels, pv, rocX)
```

@@

mp = model.performance(performance.at.1se, coefs, folds, splits )
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot
ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling(full_width = F)

rm(mp,most.stable)
```

@@

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
rm(no.hgd.plots,performance.at.1se,coefs,mp)
```

@@

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
rm(nolgd.plots,performance.at.1se,coefs,mp)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p
ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
rm(p)
```

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots,performance.at.1se,models,cvs,labels)

@@

p = ggplot(features, aes(label,hazards, fill=coef)) + geom_col() + 
  scale_fill_gradient2() + coord_flip() + theme_classic() + 
  theme(axis.ticks.y = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), legend.position = 'none') + labs(x='', y='', title='Effects') 
ggsave('plots/effects.png', plot=p, width = 3, height=1.5, units='in')
rm(p)

cvRR = features
qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
rm(qq,p)

@@

p = globalCN.plot + geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
rm(p)

p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
rm(p)
```

@@

file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs
rm(nzcoefs)

@@

p = ggplot(dplyr::select(pg.samp, Status, Samplename, RR), aes(reorder(Samplename, -RR), 1, fill=RR)) + geom_tile() + 
  scale_fill_gradient2(midpoint = 5, low = scales::muted('blue'), high = scales::muted('red')) + 
  scale_y_continuous(expand=c(0,0)) + 
  coord_flip() + labs(y='', x='') + theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.position = 'none') 
ggsave('plots/per_sample_RR.png', plot = p, width=0.5, height=3, units='in')
rm(p)

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE')) %>%
  left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))

pg.samp %>% dplyr::select(Status, -Pathology, Prediction) %>% dplyr::rename('Probability' = 'Prediction') %>%
  dplyr::mutate(Status = recode_factor(Status, 'NP' = 'Non-Progressor', 'P' = 'Progressor')) %>%
  write_tsv('plots/loo_all_samples.tsv')

performance.at.1se = enframe(performance.at.1se, value='performance.at.1se')#names(performance.at.1se) = pg.samp$Samplename
performance.at.1se$name = unique(pg.samp$Patient)
p = ggplot(performance.at.1se, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  #geom_label(data=filter(performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
rm(p,performance.at.1se,plots,coefs,fits)
```

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
  rm(plots,coefs,nzcoefs,fits)
} else {
  stop(paste("No file", file, "found"))
}

@@

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
rm(p, perf,performance.at.1se)

@@

pg.sampNOHGD %>% dplyr::select(Status, -Pathology, Prediction) %>% dplyr::rename('Probability' = 'Prediction') %>%
  dplyr::mutate(Status = recode_factor(Status, 'NP' = 'Non-Progressor', 'P' = 'Progressor')) %>%
  write_tsv('plots/loo_noHGD_samples.tsv')

rocNOHGD = pROC::roc(Status ~ Prediction, data = pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T, 0.3)

@@

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)

rm(pg.sampNOHGD)
```

@@

```{r loonolgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
  rm(plots,performance.at.1se,coefs,nzcoefs,fits)
} else {
  stop(paste("No file", file, "found"))
}

@@

ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)

rm(pg.sampNOLGD)
```

@@

@@

blues = RColorBrewer::brewer.pal(5,'Greys')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
rm(insilico.rocs, insilico.perf, blues, i.thresholds, df, p)
```

@@

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD only', incS = T), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}

tmp = pg.samp %>% dplyr::select(Patient, Status, Endoscopy.Year, PID, Prediction) %>%
  arrange(Patient, Endoscopy.Year, PID)

pout = tibble()
for (pt in unique(pg.samp$Patient)) {
  pout = pout %>% bind_rows( tmp %>% filter(Patient == pt) %>% group_by(Endoscopy.Year, PID) %>% dplyr::mutate(Endoscopy = dplyr::group_indices()) )
}

pout %>% dplyr::select(-Endoscopy.Year, -PID) %>% dplyr::rename('Probability' = 'Prediction') %>% 
  dplyr::mutate(Status = recode_factor(Status, 'NP'='Non-Progressor', 'P'='Progressor')) %>% 
  write_tsv('plots/aggregate_rocs.tsv')

pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% 
  dplyr::summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID) )

@@

multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo), threshold = 0.3) + labs(title='Endoscopy prediction ROCs')

rm(pendo, pendo.rand)
```

@@

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
rocPtM$model = 'Max per Patient'
#pROC::coords(rocPtM, 'best', transpose=T)
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
